---
title: "Homework Assignment 01"
author: "Juan Cruz Ferreyra"
date: "`r Sys.Date()`"
output:
  pdf_document:
    df_print: kable
---

&nbsp;

&nbsp;

## Section 1

**a. Create two vectors named v1 and v2, where v1 is the sequence of integers from 2 to 6, and v2 is the sequence of integers from 5 to 9.**

```{r}
# Create vector for the sequence of integers from 2 to 6
v1 <- 2:6
v1
```

```{r}
# Create vector for the sequence of integers from 5 to 9
v2 <- 5:9
v2
```

**b. What is v2 minus v1?**

```{r}
# Perform difference of vectors
v2 - v1
```

**c. What is the inner product of v1 and v2?**

```{r}
# Calculate the inner product of the two vectors
v1 %*% v2
```

**d. Replace the elements in v1+v2 that are greater than 10 with the number 0. Show that vector.**

```{r}
# Perform addition of vectors
v3 <- v1 + v2

# Replace values greater than 10
v3[v3 > 10] <- 0
v3
```

***

&nbsp;

&nbsp;

## Section 2

**a. Create a 5 by 5 matrix with the numbers 1 to 25 as its elements, and call it m1.**

```{r}
# Create a matrix for the sequence of integers from 1 to 25
m1 <- matrix(1:25, 5, 5)
m1
```

**b. What is m1 times v1?**

```{r}
# Perform matrix-vector multiplication
m1 %*% v1
```

**c. What is v1 times m1?**

```{r}
# Perform vector-matrix multiplication
v1 %*% m1
```

**d. What is m1 times the transpose of m1?**

```{r}
# Perform matrix-matrix multiplication
m1 %*% t(m1)
```

***

&nbsp;

&nbsp;

## Section 3

**a. Create a date frame with at least five rows and three columns. The first variable (column) should be dates, the second variable should be strings (characters), and the third variable should be numbers. Name each variable something appropriate and short.**

```{r}
# Create individual columns data
dates <- as.Date(c(rep("2024-09-09", 3), rep("2024-09-16", 3), rep("2024-09-23", 3)))
fruits <- rep(c("bananas", "apples", "pears"), 3)
stock <- sample(15:30, 9, replace = TRUE)

# Create a dataframe with the individual columns data
df <- data.frame(Date = dates, Fruit = fruits, Stock = stock)
df
```

**b. Use str() to show that your data frame is appropriately structured.**

```{r}
# Show dataframe structure
str(df)
```

**c. Save it as a csv file, and then reload the data from the csv file.**

```{r}
# Set directory variables
main_path <- paste0(getwd(), "/..")
file_path <- paste0(main_path, "/data/homework_assignment_01/df.csv")

# Create the directory if it doesn't exist
dir.create(dirname(file_path), recursive = TRUE, showWarnings = FALSE)

# Save df as a csv file
write.csv(df, file_path, row.names = FALSE)
```

```{r}
# Reload df from recently created csv file
reloaded_df <- read.table(file_path, header=TRUE, sep=",", stringsAsFactors=FALSE)
reloaded_df
```

**d. Create a new data frame that is just a subset of your data: the first, third, and last rows, and the first two variables.**

```{r}
# Subset dataframe
df2 <- df[c(1,3,nrow(df)), 1:2]
df2
```

**e. Replace all the even numbers in the original data frame with 0.**

```{r}
# Replace even numbers in the Stock column
is_even <- df["Stock"] %% 2 == 0
df["Stock"][is_even] <- 0
df
```

**f. Create a list with v1, v2, m1, and your data frame. Give all the items in that list a name. Now pick out the third item’s second item.**

```{r}
# Create a named list to store our data
data_list <- list(
  "v1" = v1,
  "v2" = v2,
  "m1" = m1,
  "df" = df
)

# Access second item in m1 (third item in the list)
data_list[[3]][2]
```

***

&nbsp;

&nbsp;

## Section 4

**a. Using latex equation notation in your .Rmd file, write out the quadratic formula, so that in your html file it looks pretty and like the version we all learned in high school. (Eg, see the box in the top right of this wikipedia page: http://en.wikipedia.org/wiki/Quadratic_equation.)**

$$ x = \frac{-b \pm \sqrt{b^2 - 4ac}}{2a} $$
 
​


























